# Reproducible Research: Peer Assessment 1
====================================

## Loading and preprocessing the data
```{r}
opts_chunk$set(echo = TRUE)
#setwd("./Assessment1")
#download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", 
#              destfile = "activity.zip")
unzip("activity.zip")
data.na <- read.csv("activity.csv", header=TRUE)
data <- na.omit(data.na)
```


## What is mean total number of steps taken per day?
```{r fig.width=7, fig.height=6}
hist(data$steps)
mymean <- mean(data$steps)
mymedian <- median(data$steps)
```
The mean for the total number of steps taken per day is **`r mymean`** 
 and the median is **`r mymedian`**.

## What is the average daily activity pattern?
```{r fig.width=7, fig.height=6}
avg.steps.interval=aggregate(steps~interval, data, mean)
plot(x = avg.steps.interval$interval,
     y = avg.steps.interval$steps,
     type = "l",
     main = "Average Daily Activity",
     sub = "",
     xlab = "Interval",
     ylab = "Number of steps" )
max.interval <- subset(avg.steps.interval, 
                   subset=steps==max(avg.steps.interval$steps), 
                   select= c(interval))
```
The **`r max.interval`** 5-minute interval, on average across all the days in the dataset, 
 contains the maximum number of steps.

## Imputing missing values
```{r}
na.count <- sum(is.na(data.na))
```
The total number of missing values are **`r na.count`**.

I impute the missing data by using the mean for that 5-minute interval.
```{r fig.width=7, fig.height=6}
library(plyr)
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
data2 <- ddply(data.na, ~ interval, transform, steps = impute.mean(steps))
data2 <- data2[order(data2$date, data2$interval), ] #plyr orders by group so we have to reorder
hist(data2$steps)
mymean2 <- mean(data2$steps)
mymedian2 <- median(data2$steps)
```
The mean for the total number of steps taken per day is **`r mymean`** 
 and the median is **`r mymedian`**. 
These values do **not** differ from the estimates from the first part of the assignment.
There was **no impact** when imputing missing data on the estimates of the total daily number of steps.

## Are there differences in activity patterns between weekdays and weekends?
```{r fig.width=7, fig.height=6}
data2$weekday <- as.factor(ifelse(weekdays(as.Date(data2$date)) %in% 
                                    c("Saturday","Sunday"),"weekend", "weekday"))
avg.steps.interval2=aggregate(steps~interval+weekday, data2, mean)

library(lattice)
xyplot(avg.steps.interval2$steps~avg.steps.interval2$interval|avg.steps.interval2$weekday, 
    type = "l",
    layout=(c(1,2)),
    main="Average Daily Activity", 
    ylab="Number of steps", xlab="interval")
```



